|

### 2023-01-28 paper.Rmd
Sat Jan 28 15:00:04 CET 2023: update-remote.sh
Sat Jan 28 16:00:04 CET 2023: update-remote.sh

>>>>

# Alternative Pre-Training Strategies

Since pre-trained thus far did not demonstrate significant benefits for DAE-GP, this chapter describes two alternative pre-training strategies that are also tested for the airfoil dataset:

## Second Generation Pre-Training

One common observation in EC-based optimization algorithms is that the largest improvements in fitness are often occurring in the first generation of the search process. Second Generation Pre-Training tries to exploit this observation by shifting the pre-training phase from the start of the algorithm into the second generation with the aim to produce a pre-training Population $\hat{P}$ that more accurately captures the properties of the search problem. The basic idea is to run basic DAE-GP for the first generation without any interference, after completing the first generation, the pre-training population $\hat{P}$ is created by sampling the DAE-LSTM model $M_1$. The pre-training model $\hat{M}$ is then trained using $\hat{P}$, after finishing DAE-GP continues running and initializes each new DAE-LSTM model $M_g$ for $g\in{2,...,g_{max}}$ using the trainable parameters $\theta_{\hat{M}}$.

## Grow-Initialized Pre-Training

The pre-training implementation in previous experiments used ramped half and half initialization [@Koza1993GeneticP] to create $\hat{P}$, this leads to 50% of the population consisting of full grown trees. An alternative method of initializing $\hat{P}$ is to use the grow initialization method where solutions are not guaranteed to be full trees. This different structural composition of solutions inside the pre-training population might be another way to improve the pre-training strategy by training on a population that more closely resembles the structural properties of futher generations population where a lower percentage of full grown trees is to be expected.

## Results

Figure \@ref(fig:altStrategies-Fitness) shows the median best fitness by generations for both alternative pre-training strategies as well as the results for standard DAE-GP and the standard Pre-Training implementation used in earlier exoeriments.

```{r altStrategies-Fitness, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/median_fitness_byGens.png")
```

Figure \@ref(fig:altStrategies-finalFitness)

```{r altStrategies-FitnessDistr, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Alternative Pre-Trategies - Airfoil", out.width="80%"}
knitr::include_graphics("./img/PT_Strategies_airfoil_2hl_150hn_fullRun/final_fit_boxplot(test_only).png")
```
